### 2017-04-05 writing/paper/sections/abstract.tex
Finish writing paper v1
Comments from Henning up to section 4

\begin{abstract}
    Two mesh patterns are coincident if they are avoided by the same set of permutations, and are
    Wilf-equivalent if they have the same number of avoiders at each length. We provide
    sufficient conditions for coincidence among mesh patterns, whilst also avoiding a longer
    classical pattern. Using these conditions we completely classify coincidences between families
    containing a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
    completely Wilf-classify equivalences of mesh patterns of length 2 whilst also avoiding the
    classical pattern 231.

>>>>

\begin{abstract} Two mesh patterns are coincident if they are avoided by the
same set of permutations, and are Wilf-equivalent if they have the same number
of avoiders of each length. We provide sufficient conditions for coincidence
among mesh patterns, whilst also avoiding a longer classical pattern. Using
these conditions we completely classify coincidences between families containing
a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
completely Wilf-classify equivalences of mesh patterns of length 2 whilst also
avoiding the classical pattern 231.

### 2017-04-05 writing/paper/sections/coincidences.tex
Comments from Henning up to section 4

Two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
\emph{containers},\ie \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
and \(p\) is a mesh pattern of length 2 in order to establish sufficient conditions for two such
sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that
\(\pi\) is the \emph{dominating pattern}.

In order to describe the rules it is useful to have a notion for inserting points, ascents, and
descents into a mesh pattern.
\begin{definition}
\label{def:ap}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a
point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying
classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
\end{cases}
\end{equation*}
While the mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
\end{aligned}
\end{equation*}
\end{definition}
In addition, we give the following definitions:
\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and
\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) to be as defined in \DefinitionRef{def:ap}. We define
the following five modifications of the mesh patterns of the same length as \(p^\boks{i,j}\).
\begin{align*}
p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
\end{align*}
\end{definition}
Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in
\(\boks{i,j}\). We allow composition of these modifications, and collect the resulting mesh patterns
in a set
\begin{equation*}
p^\boksall{i,j}=\left\{
p^\boks{i,j},
p^\boksright{i,j},
% p^\boksrightup{i,j},
% p^\boksrightdown{i,j},
p^\boksleft{i,j},
% p^\boksleftup{i,j},
% p^\boksleftdown{i,j},
p^\boksup{i,j},
p^\boksdown{i,j}\right\}
\end{equation*}
See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.

@@

We now attempt to fully classify coincidences in families characterised by avoidance
of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).

@@

We know that there are a total of \(512\) mesh-sets for each underlying classical
pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}\footnote{
The authors use the Simultaneous Shading Lemma, a closure result and one worked out special case.
} the number of coincidence classes can be reduced to \(220\).

@@

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
    \end{enumerate}
\end{proposition}
In order to prove this proposition we must first make the following note.

\begin{note}
    \label{not:downcmesh}
    Let \(R^\prime \subseteq R\). Then any occurrence of \((\tau, R)\) in a permutation
    is an occurrence of \((\tau,R^\prime)\).
\end{note}

    Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
    containing an occurrence of \(m_1\). Consider placing a point in the
    region corresponding to the box \((a,b)\), regardless of where in this
    region we place the point by condition~\ref{prop:dom1:cont} of the
    Proposition we create an occurrence of \(\pi\), therefore there can be no
    points in this region, which could have been represented in the mesh set
    \(R_1\) by adding the box \((a,b)\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

@@

\begin{lemma}
    \label{lem:incdec}
    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
    in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
    \(\pi \preceq p^{\boks{i,j}_a}\)\\(\(\pi \preceq p^{\boks{i,j}_d}\))
    ,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
    corresponding to the box \((a,b)\) can only contain an decreasing
    (increasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
is an occurrence of \(m_2\). This shows that our two example patterns are coincident.

@@

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item   \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{i,j}_a}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq p^{\boks{i,j}_d}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
    By \NoteRef{not:downcmesh} any of these occurrences is also an occurrence
    of \(m_1\). This proves that every occurrence of \(m_2\) is also an
    occurrence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    We will consider taking the first branch of every choice.
    Now consider a permutation \(\omega^\prime \in \av{\pi}\). Suppose
    \(\omega^\prime\) contains \(m_1\) and consider the region corresponding to
    \((a,b)\) in \(R_1\).

    If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
    of \(m_2\).

    If the region is non-empty, then by \LemmaRef{lem:incdec} and
    condition~\ref{prop:dom2:condc} of the proposition it must contain
    a decreasing subsequence.
    We can choose the topmost point in the region to replace the corresponding
    point in the mesh pattern and the points from the subsequence are now in
    the box southeast of the point. The other conditions allow this to be done
    without points being present in regions that were shaded. Hence there are no
    points in the region corresponding to the box \((a,b)\) in the mesh
    pattern, and therefore we can shade this region. This implies that every
    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

By taking the First Dominating Pattern Rule and the Second Dominating Pattern
Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
are fully explained, obtaining 39 coincidence classes of mesh patterns, confirming experimental
observations.

@@

For example the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are experimentally coincident. This coincidence is not explained by our rules,
it is necessary to attempt to capture these coincidences by establishing more
rules.

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty. Consider choosing the rightmost point in
region.

@@

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} One of \((\sigma,R_1)^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
    By \NoteRef{not:downcmesh},
     \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.

    Now consider a permutation \(\varrho\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
    is empty then we have an occurrence of \(m_2\). If the region is non-empty
    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
    a direction such that there exists an occurrence of a mesh pattern of
    length one longer than \(m_1\) in this position. This mesh pattern is
    coincident with another mesh pattern that contains an
    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

By application of this rule we can reduce the number of classes in
\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the 
56 classes observed through experimentation

@@

In order to prove this coincidence we will proceed by mathematical
induction on the number of points in region corresponding to the middle box.
We call this number \(n\).
\begin{description}
    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
        box if it contains no points.
    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.
    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, this gives the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done. If this box contains
        any points then we gain some extra shading on the mesh pattern
        as any points in these boxes would create an occurrence of the dominating pattern \(\perm{2,3,1}\)
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.
\end{description}
By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
\(m_1\) so the two patterns are coincident.
This argument applies to another two pairs of classes. Therefore in
total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) coincidence
classes.

>>>>

Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
\emph{containers}, \ie \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
and \(p\) is a mesh pattern of length 2 and establish sufficient conditions for two such
sets to be coincident. The classical pattern \(\pi\) will be fixed and called
the \emph{dominating pattern}.

In order to describe the rules it is useful to have a notion for inserting points, ascents, and
descents into a mesh pattern.
\begin{definition}
\label{def:ap}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a
point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying
classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
\end{cases}
\end{equation*}
While the mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
\end{aligned}
\end{equation*}
\end{definition}
In addition, we give the following definitions:

\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and
\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) is as defined in \DefinitionRef{def:ap}. We define
the following four modifications of \(p^\boks{i,j}\).
\begin{align*}
p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
\end{align*}
\end{definition}
Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in
\(\boks{i,j}\). We collect the resulting mesh patterns in a set
\begin{equation*}
p^\boksall{i,j}=\left\{
p^\boks{i,j},
p^\boksright{i,j},
% p^\boksrightup{i,j},
% p^\boksrightdown{i,j},
p^\boksleft{i,j},
% p^\boksleftup{i,j},
% p^\boksleftdown{i,j},
p^\boksup{i,j},
p^\boksdown{i,j}\right\}
\end{equation*}
We allow composition of these modifications. % TODO Are they ever used?
See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.

@@

We now attempt to fully classify coincidences in families characterised by avoidance
of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).
% TODO There is some inconsistency in when you put the set brackets into Av

@@

We know that there are a total of \(512\) mesh-sets for each underlying classical
pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}
the number of coincidence classes can be reduced to \(220\).

@@

    \begin{enumerate}
        \item The symmetric difference, \(R_1 \triangle R_2\), is \(\{(a,b)\}\)
        \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
    \end{enumerate}
\end{proposition}

@@

In order to prove the proposition we must first make the following note.

\begin{note}
    \label{not:downcmesh}
    Let \(R_1 \subseteq R_2\). Then any occurrence of \((\tau, R_2)\) in a permutation
    is an occurrence of \((\tau,R_1)\).
\end{note}

    Now we consider a permutation \(\omega \in \av{\pi}\),
    containing an occurrence of \(m_1\). Consider placing a point in the
    region corresponding to the box \((a,b)\), regardless of where in this
    region we place the point by condition~\ref{prop:dom1:cont} of the
    proposition we create an occurrence of \(\pi\). Therefore there can be no
    points in this region, which implies this is an occurrence of $m_2$. Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

\begin{lemma}
    \label{lem:incdec}
    Let \(m =(\sigma, R)\) be a mesh pattern, where the box \((a,b)\) is not
    in \(R\), and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
    \(\pi \preceq p^{\boks{i,j}_a}\)\\(\(\pi \preceq p^{\boks{i,j}_d}\)),
    then in any occurrence of \(m\) in a permutation \(\varrho\), the region
    corresponding to the box \((a,b)\) can only contain an decreasing
    (increasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
we know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
is an occurrence of \(m_2\). This shows that our two example patterns are coincident.

@@

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item   \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{i,j}_a}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq p^{\boks{i,j}_d}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
    implies and occurrence of \(m_2\).
    We consider taking the first branch of every choice.
    Now consider a permutation \(\omega \in \av{\pi}\). Suppose
    \(\omega\) contains \(m_1\) and consider the region corresponding to
    \((a,b)\) in \(R_1\).If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
    of \(m_2\).
    If the region is non-empty, then by \LemmaRef{lem:incdec} and
    condition~\ref{prop:dom2:condc} of the proposition it must contain
    a decreasing subsequence.
    We can choose the topmost point in the region to replace the corresponding
    point in the mesh pattern and the points from the subsequence are now in
    the box southeast of the point. The other conditions allow this to be done
    without points being present in regions that were shaded. Hence there are no
    points in the region corresponding to the box \((a,b)\) in the mesh
    pattern, and therefore we can shade this region. This implies that every
    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

By applying both the First Dominating Pattern Rule and the Second Dominating
Pattern Rule fully explains coincidences of classes of the form
\(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\), obtaining 39 coincidence classes of
mesh patterns, confirming experimental observations.

@@

For example the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are experimentally coincident.

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty, adn consider choosing the rightmost
point in the region.

@@

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \((\sigma,R_1)^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    \NoteRef{not:downcmesh}, implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.

    Now consider a permutation \(\sigma\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
    is empty then we have an occurrence of \(m_2\). If the region is non-empty
    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
    a pattern in \((\sigma,R_1)^\boksall{a,b}\) such that there exists an occurrence of a mesh pattern of
    length one longer than \(m_1\) in this position. This mesh pattern is
    coincident with another mesh pattern that contains an
    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

By application of this rule we can reduce the number of classes in
\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the
56 classes observed through experimentation.

@@

In order to prove this coincidence we will proceed by mathematical
induction on the number of points in region corresponding to the middle box.
We call this number \(n\).
\begin{description}
    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
        box if it contains no points.
    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.
    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, which gives the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done. If this box contains
        any points then we gain some extra shading on the mesh pattern
        as any points in these boxes would create an occurrence of the dominating pattern \(\perm{2,3,1}\)
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.
\end{description}
By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
\(m_1\) so the two patterns are coincident.
This argument applies to another two pairs of classes. Therefore in
total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) coincidence
classes.

### 2017-04-05 writing/paper/sections/introduction.tex
Make Changes suggest by Michael
Comments from Henning up to section 4

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions.
\textcite{babstein2000} considered \emph{vincular} patterns (also known as
\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
pattern must also be adjacent in the permutation. \textcite{MR2652101} look at
classes of pattern where both columns and rows can be shaded, these are called
\emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by
\textcite{MR2264071} to establish necessary conditions for a Schubert variety to
be Gorenstein. These definitions are subsumed under the definition of \emph{mesh
patterns}, introduced by \textcite{journals/combinatorics/BrandenC11} to capture
explicit expansions for certain permutation statistics.

@@

In this work we aim to establish some ground in this field by computing
coincidences and Wilf-classes and calculating some of the enumerations of
avoiders of a mesh pattern of length 2 and a classical pattern of length 3. We
begin by establishing coincidences between mesh patterns of length 2 while
avoiding a classical pattern of length 3, this is used to establish sufficient
conditions for coincidence. We then establish Wilf-equivalence classes of these
coincidence classes who avoid the classical pattern 231.

>>>>

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions.
\textcite{babstein2000} considered \emph{vincular} patterns (also known as
\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
look at classes of pattern where entries can be required to be consecutive in
value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
patterns were studied by \textcite{MR2264071} to establish necessary conditions
for a Schubert variety to be Gorenstein. These definitions are subsumed under
the definition of \emph{mesh patterns}, introduced by
\textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for
certain permutation statistics.

@@

In this work we aim to establish some ground in this field by computing
coincidences and Wilf-classes and calculating some of the enumerations of
avoiders of a mesh pattern of length 2 and a classical pattern of length 3. We
begin by establishing coincidences between mesh patterns of length 2 while
avoiding a classical pattern of length 3; this is used to establish sufficient
conditions for coincidence. We then establish Wilf-equivalence classes of these
coincidence classes where the classical pattern 231.
